---
title: "Untitled"
format: html
editor: visual
---


# Model with resource uptake
## initial model
```{r}
library(ggplot2)
library(tidyverse)

# Main function to create and run the model
resource_model <- function(r, t) {
  # Initialize resource pools
  initial_N <- 100
  initial_P <- 100
  N_pool <- initial_N
  P_pool <- initial_P
  tr <- 12
  community_richness <- tr/r
  
  # Create species characteristics
  species_traits <- data.frame(
    species = 1:r,
    xn = runif(r, min=0.01, max=1),
    yp = runif(r, min=0.01, max=1),
    growth_time = runif(r, min=10, max=20)
  )
  
  # Calculate derived traits
  species_traits$npr <- species_traits$xn / species_traits$yp
  species_traits$k_max <- sqrt(species_traits$xn^2 + species_traits$yp^2)
  
  # Initialize matrices and vectors to track changes
  biomass_time <- matrix(0, nrow=t, ncol=r)
  total_biomass <- numeric(t)
  N_available <- numeric(t)
  P_available <- numeric(t)
  
  # Simulation loop
  for(time in 1:t) {
    # Store resource availability
    N_available[time] <- N_pool
    P_available[time] <- P_pool
    
    # Calculate current absorption rates for each species
    current_k <- sapply(1:r, function(i) {
      if(time >= species_traits$growth_time[i]) {
        species_traits$k_max[i]
      } else {
        species_traits$k_max[i] * (time/species_traits$growth_time[i])
      }
    })
    
    # Calculate resource uptake and biomass production for each species
    for(i in 1:r) {
      # Calculate potential resource uptake
      N_uptake <- current_k[i] * species_traits$xn[i] * community_richness
      P_uptake <- current_k[i] * species_traits$yp[i] * community_richness
      
      # Adjust uptake based on available resources
      N_limitation <- N_uptake/N_pool
      P_limitation <- P_uptake/P_pool
      
      # Use the most limiting resource to determine actual uptake
      limitation_factor <- min(1, 1/exp(max(N_limitation, P_limitation)))
      
      # Calculate actual resource uptake
      actual_N_uptake <- N_uptake * limitation_factor
      actual_P_uptake <- P_uptake * limitation_factor
      
      # Update resource pools
      N_pool <- N_pool - actual_N_uptake
      P_pool <- P_pool - actual_P_uptake
      
      # Calculate biomass production
      new_biomass <- (actual_N_uptake + actual_P_uptake)
      biomass_time[time, i] <- if(time > 1) biomass_time[time-1, i] + new_biomass else new_biomass
    }
    
    # Calculate total community biomass for this time step
    total_biomass[time] <- sum(biomass_time[time,])
    
    # Stop if resources are depleted
    if(N_pool <= 0 || P_pool <= 0) {
      if(time < t) {
        # Fill remaining time steps with final values
        for(remain_t in (time+1):t) {
          biomass_time[remain_t,] <- biomass_time[time,]
          total_biomass[remain_t] <- total_biomass[time]
          N_available[remain_t] <- N_available[time]
          P_available[remain_t] <- P_available[time]
        }
      }
      break
    }
  }
  
  # Create visualizations
  # 1. Resource utilization coordinate system
  plot1 <- ggplot(species_traits, aes(x=xn, y=yp)) +
    geom_point(aes(color=k_max, size=npr)) +
    labs(title="Species Resource Requirements",
         x="Nitrogen Requirement (xn)",
         y="Phosphorus Requirement (yp)",
         color="Absorption Rate (k)",
         size="N:P Ratio") +
    theme_minimal()
  
  # 2. Time vs Biomass
  biomass_df <- as.data.frame(biomass_time)
  biomass_df$time <- 1:t
  biomass_long <- pivot_longer(biomass_df, 
                              cols=-time, 
                              names_to="species", 
                              values_to="biomass")
  
  plot2 <- ggplot(biomass_long, aes(x=time, y=biomass, color=species)) +
    geom_line() +
    geom_line(data=data.frame(time=1:t, biomass=total_biomass), 
              aes(x=time, y=biomass), color="black", size=1.2) +
    labs(title="Biomass Over Time",
         x="Time",
         y="Biomass") +
    theme_minimal()
  
  # 3. k vs Final Biomass
  plot3 <- ggplot(species_traits, aes(x=k_max, y=biomass_time[t,])) +
    geom_point() +
    geom_smooth(method = "lm", se = TRUE) +
    labs(title="Final Biomass vs Absorption Rate",
         x="Absorption Rate (k)",
         y="Final Biomass") +
    theme_minimal()
  
  # 4. npr vs Final Biomass
  plot4 <- ggplot(species_traits, aes(x=npr, y=biomass_time[t,])) +
    geom_point() +
    labs(title="Final Biomass vs N:P Ratio",
         x="N:P Ratio",
         y="Final Biomass") +
    theme_minimal()
  
  # Return results and plots
  return(list(
    species_traits = species_traits,
    biomass_time = biomass_time,
    total_biomass = total_biomass,
    N_available = N_available,
    P_available = P_available,
    plots = list(
      resource_plot = plot1,
      time_plot = plot2,
      k_plot = plot3,
      npr_plot = plot4
    )
  ))
}

# Example usage:
results <- resource_model(r=12, t=50)
# View plots
results$plots$resource_plot
results$plots$time_plot
results$plots$k_plot
results$plots$npr_plot
```
## richness
```{r}
library(ggplot2)
library(tidyverse)
library(patchwork)

compare_richness <- function(richness_levels = c(1, 4, 8, 12), num_runs = 10, t = 50) {
  # Create empty lists to store results
  biomass_results <- list()
  nutrient_results <- list()
  
  # Run simulations for each richness level and run
  for(r in richness_levels) {
    for(run in 1:num_runs) {
      sim_result <- resource_model(r = r, t = t)
      
      # Store biomass data
      biomass_results[[length(biomass_results) + 1]] <- data.frame(
        richness = r,
        run = run,
        time = 1:t,
        biomass = sim_result$total_biomass
      )
      
      # Store final nutrient data
      nutrient_results[[length(nutrient_results) + 1]] <- data.frame(
        richness = r,
        run = run,
        N_remaining = tail(sim_result$N_available, 1),
        P_remaining = tail(sim_result$P_available, 1)
      )
    }
  }
  
  # Combine results
  biomass_df <- do.call(rbind, biomass_results)
  nutrient_df <- do.call(rbind, nutrient_results)
  
  # Figure 1: Average biomass over time with confidence intervals
  avg_biomass <- biomass_df %>%
    group_by(richness, time) %>%
    summarize(
      mean_biomass = mean(biomass),
      sd_biomass = sd(biomass),
      ci = 1.96 * sd_biomass / sqrt(n()),
      upper = mean_biomass + ci,
      lower = mean_biomass - ci,
      .groups = 'drop'
    )
  
  plot1 <- ggplot(avg_biomass, aes(x = time, y = mean_biomass, color = factor(richness))) +
    geom_line(size = 1) +
    geom_ribbon(aes(ymin = lower, ymax = upper, fill = factor(richness)), alpha = 0.2) +
    labs(title = "Average Community Biomass Over Time",
         x = "Time",
         y = "Community Biomass",
         color = "Species Richness",
         fill = "Species Richness") +
    theme_minimal()
  
  # Figure 2: Final biomass boxplot
  final_biomass <- biomass_df %>%
    group_by(richness, run) %>%
    slice_max(time, n = 1)
  
  plot2 <- ggplot(final_biomass, aes(x = factor(richness), y = biomass)) +
    geom_boxplot(aes(fill = factor(richness))) +
    labs(title = "Final Community Biomass by Species Richness",
         x = "Species Richness",
         y = "Final Biomass",
         fill = "Species Richness") +
    theme_minimal()
  
  # Figure 3: Remaining nutrients boxplot
  nutrient_long <- nutrient_df %>%
    pivot_longer(cols = c(N_remaining, P_remaining),
                names_to = "nutrient",
                values_to = "amount")
  
  plot3 <- ggplot(nutrient_long, aes(x = factor(richness), y = amount)) +
    geom_boxplot(aes(fill = nutrient)) +
    labs(title = "Remaining Nutrients by Species Richness",
         x = "Species Richness",
         y = "Amount Remaining",
         fill = "Nutrient") +
    theme_minimal()
  
  # Combine plots
  combined_plots <- plot1 / plot2 / plot3 +
    plot_layout(heights = c(2, 1, 1)) +
    plot_annotation(title = "Community Dynamics Across Species Richness Levels")
  
  return(list(
    biomass_data = biomass_df,
    nutrient_data = nutrient_df,
    plots = list(
      biomass_time = plot1,
      final_biomass = plot2,
      remaining_nutrients = plot3,
      combined = combined_plots
    )
  ))
}

# Run the analysis
set.seed(123)
results <- compare_richness()

# Display plots
print(results$plots$combined)

# Print summary statistics
final_stats <- results$biomass_data %>%
  group_by(richness, run) %>%
  slice_max(time, n = 1) %>%
  group_by(richness) %>%
  summarize(
    mean_biomass = mean(biomass),
    sd_biomass = sd(biomass),
    se_biomass = sd_biomass/sqrt(n())
  )

print("Summary Statistics for Final Biomass:")
print(final_stats)


```



## Formula
Based on our model, the biomass formula for each species can be described as follows:

At each time step t, for species i:

Biomass(t,i) = Biomass(t-1,i) + new_biomass(t,i)

Where new_biomass(t,i) = actual_N_uptake + actual_P_uptake

And:
actual_N_uptake = current_k(t,i) * xn(i) * limitation_factor
actual_P_uptake = current_k(t,i) * yp(i) * limitation_factor

Where:
current_k(t,i) = {
    k_max(i) * (t/growth_time(i))  if t < growth_time(i)
    k_max(i)                       if t ≥ growth_time(i)
}

k_max(i) = sqrt(xn(i)^2 + yp(i)^2)

limitation_factor = min(1, 1/exp(max(N_limitation, P_limitation)))
N_limitation = (current_k(t,i) * xn(i))/N_pool
P_limitation = (current_k(t,i) * yp(i))/P_pool

Variable meanings:

Biomass(t,i): Total accumulated biomass of species i at time t
xn(i): Nitrogen requirement coefficient for species i (random value between 0-1)
yp(i): Phosphorus requirement coefficient for species i (random value between 0-1)
k_max(i): Maximum absorption rate for species i
growth_time(i): Time needed to reach maximum absorption rate (random value between 1-10)
current_k(t,i): Current absorption rate at time t for species i
N_pool: Available nitrogen in the environment
P_pool: Available phosphorus in the environment
limitation_factor: Factor that adjusts resource uptake based on resource availability
N_limitation: Nitrogen limitation ratio
P_limitation: Phosphorus limitation ratio

This formula captures:
1.Linear growth of absorption rate until reaching maximum
2.Resource limitation effects
3.Biomass accumulation over time
4.Resource-dependent growth

# Model with resource uptake and species interactions
## niche-biomass relationship
```{r}
# Libraries
library(ggplot2)
library(patchwork)

# Function to calculate niche overlap and ri values
calculate_niche_interactions <- function(species_traits) {
  num_species <- nrow(species_traits)
  
  # Initialize overlap matrix
  overlap_matrix <- matrix(0, nrow = num_species, ncol = num_species)
  
  # Calculate pairwise niche overlap (i) between each pair of species
  for(i in 1:num_species) {
    for(j in 1:num_species) {
      if(i != j) {
        # Calculate range of potential overlap
        x_min <- min(species_traits$np[i] - species_traits$nb[i]/2, 
                    species_traits$np[j] - species_traits$nb[j]/2)
        x_max <- max(species_traits$np[i] + species_traits$nb[i]/2, 
                    species_traits$np[j] + species_traits$nb[j]/2)
        x <- seq(x_min, x_max, length.out = 1000)
        
        # Calculate Gaussian curves ONLY for overlap calculation
        y1 <- exp(-(x - species_traits$np[i])^2 / (2 * (species_traits$nb[i]/4)^2))
        y2 <- exp(-(x - species_traits$np[j])^2 / (2 * (species_traits$nb[j]/4)^2))
        
        # Calculate overlap area (i)
        overlap_matrix[i,j] <- sum(pmin(y1, y2)) * (x_max - x_min) / 1000
      }
    }
  }
  
  # Calculate total i for each species
  i_values <- rowSums(overlap_matrix)
  
  # Calculate ri values
  ri_values <- i_values + 1
  
  return(list(
    overlap_matrix = overlap_matrix,
    i_values = i_values,
    ri_values = ri_values
  ))
}

# Function to calculate ni based on ri and m
calculate_ni <- function(ri, m) {
  if(m == 0) {
    return(1)  # When m=0, ni=1 regardless of ri
  } else if(m == 1) {
    return(ri)  # Facilitation
  } else if(m == -1) {
    return(1/ri)  # Competition
  } else {
    stop("m must be -1, 0, or 1")
  }
}

# Basic Resource Competition Model (without niche)
basic_resource_model <- function(r, t) {
  # Initialize resource pools
  initial_N <- 100
  initial_P <- 100
  N_pool <- initial_N
  P_pool <- initial_P
  tr <- 12
  community_richness <- tr/r
  
  # Create species characteristics
  species_traits <- data.frame(
    species = 1:r,
    xn = runif(r, min=0.1, max=1),
    yp = runif(r, min=0.1, max=1),
    growth_time = runif(r, min=10, max=20)
  )
  
  # Calculate base kmax
  species_traits$k_max <- sqrt(species_traits$xn^2 + species_traits$yp^2)
  
  # Initialize tracking matrices
  biomass_time <- matrix(0, nrow=t, ncol=r)
  total_biomass <- numeric(t)
  N_available <- numeric(t)
  P_available <- numeric(t)
  
  # Simulation loop
  for(time in 1:t) {
    N_available[time] <- N_pool
    P_available[time] <- P_pool
    
    if(N_pool > 0 && P_pool > 0) {
      # Calculate current acquisition rates
      current_k <- sapply(1:r, function(i) {
        if(time < species_traits$growth_time[i]) {
          species_traits$k_max[i] * (1 / (1 + exp(-5 * (time/species_traits$growth_time[i] - 0.5))))
        } else {
          species_traits$k_max[i]
        }
      })
      
      for(i in 1:r) {
        # Calculate potential uptake
        N_uptake <- current_k[i] * species_traits$xn[i] * community_richness
        P_uptake <- current_k[i] * species_traits$yp[i] * community_richness
        
        # Resource limitation
        N_limitation <- N_uptake/N_pool
        P_limitation <- P_uptake/P_pool
        limitation_factor <- min(1, 1/exp(max(N_limitation, P_limitation)))
        
        # Actual resource uptake
        actual_N_uptake <- N_uptake * limitation_factor
        actual_P_uptake <- P_uptake * limitation_factor
        
        # Update resource pools
        N_pool <- max(0, N_pool - actual_N_uptake)
        P_pool <- max(0, P_pool - actual_P_uptake)
        
        # Calculate new biomass increment
        new_biomass <- actual_N_uptake + actual_P_uptake
        
        # Accumulate biomass
        if(time == 1) {
          biomass_time[time, i] <- new_biomass
        } else {
          biomass_time[time, i] <- biomass_time[time-1, i] + new_biomass
        }
      }
    } else {
      if(time > 1) {
        biomass_time[time,] <- biomass_time[time-1,]
      }
    }
    
    total_biomass[time] <- sum(biomass_time[time,])
  }
  
  return(list(
    species_traits = species_traits,
    biomass_time = biomass_time,
    total_biomass = total_biomass,
    N_available = N_available,
    P_available = P_available
  ))
}

# Niche-based Resource Model
resource_model <- function(r, t, m) {
  # Initialize resource pools
  initial_N <- 100
  initial_P <- 100
  N_pool <- initial_N
  P_pool <- initial_P
  tr <- 12
  community_richness <- tr/r
  
  # Create species characteristics
  species_traits <- data.frame(
    species = 1:r,
    xn = runif(r, min=0.1, max=0.3),
    yp = runif(r, min=0.1, max=0.3),
    growth_time = runif(r, min=1, max=5),
    np = runif(r, min=0, max=100),
    nb = runif(r, min=5, max=20)
  )
  
  # Calculate base kmax
  species_traits$k_max <- sqrt(species_traits$xn^2 + species_traits$yp^2)
  
  # If m != 0, modify kmax based on niche overlap
  if(m != 0) {
    niche_results <- calculate_niche_interactions(species_traits)
    species_traits$ri <- niche_results$ri_values
    species_traits$ni <- sapply(species_traits$ri, function(ri) calculate_ni(ri, m))
    species_traits$k_max <- species_traits$k_max * species_traits$ni
  } else {
    # When m=0, no niche modification needed
    species_traits$ri <- rep(1, r)
    species_traits$ni <- rep(1, r)
  }
  
  # Initialize tracking matrices
  biomass_time <- matrix(0, nrow=t, ncol=r)
  total_biomass <- numeric(t)
  N_available <- numeric(t)
  P_available <- numeric(t)
  
  # Simulation loop
  for(time in 1:t) {
    N_available[time] <- N_pool
    P_available[time] <- P_pool
    
    if(N_pool > 0 && P_pool > 0) {
      # Calculate current acquisition rates
      current_k <- sapply(1:r, function(i) {
        if(time < species_traits$growth_time[i]) {
          species_traits$k_max[i] * (1 / (1 + exp(-5 * (time/species_traits$growth_time[i] - 0.5))))
        } else {
          species_traits$k_max[i]
        }
      })
      
      for(i in 1:r) {
        # Calculate potential uptake
        N_uptake <- current_k[i] * species_traits$xn[i] * community_richness
        P_uptake <- current_k[i] * species_traits$yp[i] * community_richness
        
        # Resource limitation
        N_limitation <- ifelse(N_pool > 0, N_uptake/N_pool, 0)
        P_limitation <- ifelse(P_pool > 0, P_uptake/P_pool, 0)
        limitation_factor <- min(1, 1/exp(max(N_limitation, P_limitation)))
        
        # Actual resource uptake
        actual_N_uptake <- N_uptake * limitation_factor
        actual_P_uptake <- P_uptake * limitation_factor
        
        # Update resource pools
        N_pool <- max(0, N_pool - actual_N_uptake)
        P_pool <- max(0, P_pool - actual_P_uptake)
        
        # Calculate new biomass increment
        new_biomass <- actual_N_uptake + actual_P_uptake
        
        # Accumulate biomass
        if(time == 1) {
          biomass_time[time, i] <- new_biomass
        } else {
          biomass_time[time, i] <- biomass_time[time-1, i] + new_biomass
        }
      }
    } else {
      if(time > 1) {
        biomass_time[time,] <- biomass_time[time-1,]
      }
    }
    
    total_biomass[time] <- sum(biomass_time[time,])
  }
  
  return(list(
    species_traits = species_traits,
    biomass_time = biomass_time,
    total_biomass = total_biomass,
    N_available = N_available,
    P_available = P_available,
    niche_overlap = if(m != 0) niche_results$overlap_matrix else matrix(0, r, r)
  ))
}

# Visualization function
create_visualizations <- function(results) {
  # Plot niche positions and breadths
  niche_data <- data.frame(
    species = factor(1:nrow(results$species_traits)),
    position = results$species_traits$np,
    breadth = results$species_traits$nb
  )
  
  fig1 <- ggplot(niche_data) +
    geom_errorbar(aes(x = species, 
                     ymin = position - breadth/2,
                     ymax = position + breadth/2,
                     color = species),
                 width = 0.2) +
    geom_point(aes(x = species, y = position, color = species), size = 3) +
    labs(title = "Species Niche Positions and Breadths",
         x = "Species",
         y = "Niche Position") +
    theme_minimal()
  
  # Biomass over time
  biomass_data <- data.frame(
    time = rep(1:nrow(results$biomass_time), ncol(results$biomass_time)),
    biomass = as.vector(results$biomass_time),
    species = factor(rep(1:ncol(results$biomass_time), 
                        each = nrow(results$biomass_time)))
  )
  
  fig2 <- ggplot(biomass_data, aes(x = time, y = biomass, color = species)) +
    geom_line(size = 1) +
    labs(title = "Species Biomass Over Time",
         x = "Time",
         y = "Biomass") +
    theme_minimal()
  
  # Resource availability
  resource_data <- data.frame(
    time = 1:length(results$N_available),
    N = results$N_available,
    P = results$P_available
  )
  
  fig3 <- ggplot(resource_data) +
    geom_line(aes(x = time, y = N, color = "N"), size = 1) +
    geom_line(aes(x = time, y = P, color = "P"), size = 1) +
    labs(title = "Resource Availability",
         x = "Time",
         y = "Amount",
         color = "Resource") +
    theme_minimal()
  
  # Combine plots
  combined_plot <- fig1 + fig2 + fig3 +
    plot_layout(ncol = 3) +
    plot_annotation(title = "Community Dynamics")
  
  return(list(
    niche_plot = fig1,
    biomass_plot = fig2,
    resource_plot = fig3,
    combined_plot = combined_plot
  ))
}

# Example usage
set.seed(123)
r <- 12  # number of species
t <- 200 # time steps

# Run models and create visualizations
basic_results <- basic_resource_model(r, t)
niche_results_m0 <- resource_model(r, t, m=0)
niche_results_m1 <- resource_model(r, t, m=1)
niche_results_m_1 <- resource_model(r, t, m=-1)

# Create visualizations
vis_m0 <- create_visualizations(niche_results_m0)
vis_m1 <- create_visualizations(niche_results_m1)
vis_m_1 <- create_visualizations(niche_results_m_1)

# Display results
print(vis_m0$combined_plot)
print(vis_m1$combined_plot)
print(vis_m_1$combined_plot)
```
### run one time
```{r}
# First run the model with m=1
# set.seed(123)
r <- 12  # number of species
t <- 100 # time steps
results <- resource_model(r, t, m=1)

# Create visualization function for these specific plots
create_specific_plots <- function(results) {
  # Figure 1: Species Niche Distribution
  niche_data <- data.frame()
  for(i in 1:nrow(results$species_traits)) {
    x <- seq(0, 100, length.out = 1000)
    y <- exp(-(x - results$species_traits$np[i])^2 / (2 * (results$species_traits$nb[i]/4)^2))
    niche_data <- rbind(niche_data, data.frame(
      x = x,
      y = y,
      species = factor(i)
    ))
  }
  
  fig1 <- ggplot(niche_data, aes(x = x, y = y, color = species)) +
    geom_line(size = 1) +
    labs(title = "Species Niche Distribution",
         x = "Niche Position",
         y = "Niche Value") +
    theme_minimal()
  
  # Figure 2: Relationship between ri and final biomass
  final_biomass <- results$biomass_time[nrow(results$biomass_time),]
  ri_biomass_data <- data.frame(
    species = factor(1:length(results$species_traits$ri)),
    ri = results$species_traits$ri,
    biomass = final_biomass
  )
  
  fig2 <- ggplot(ri_biomass_data, aes(x = ri, y = biomass)) +
    geom_point(aes(color = species), size = 3) +
    labs(title = "Relationship between ri and Final Biomass",
         x = "ri Value",
         y = "Final Biomass") +
    theme_minimal()
  
  # Figure 3: Species biomass over time
  biomass_data <- data.frame(
    time = rep(1:nrow(results$biomass_time), ncol(results$biomass_time)),
    biomass = as.vector(results$biomass_time),
    species = factor(rep(1:ncol(results$biomass_time), 
                        each = nrow(results$biomass_time)))
  )
  
  fig3 <- ggplot(biomass_data, aes(x = time, y = biomass, color = species)) +
    geom_line(size = 1) +
    labs(title = "Species Biomass Over Time",
         x = "Time",
         y = "Biomass") +
    theme_minimal()
  
  # Figure 4: Total community biomass over time
  community_data <- data.frame(
    time = 1:length(results$total_biomass),
    total_biomass = results$total_biomass
  )
  
  fig4 <- ggplot(community_data, aes(x = time, y = total_biomass)) +
    geom_line(size = 1, color = "blue") +
    labs(title = "Community Biomass Over Time",
         x = "Time",
         y = "Total Biomass") +
    theme_minimal()
  
  # Combine all plots
  combined_plot <- (fig1 + fig2) / (fig3 + fig4) +
    plot_annotation(
      title = "Community Dynamics with Niche Interactions (m=1)",
      theme = theme_minimal()
    )
  
  return(list(
    niche_plot = fig1,
    ri_biomass_plot = fig2,
    species_biomass_plot = fig3,
    community_biomass_plot = fig4,
    combined_plot = combined_plot
  ))
}

# Create and display the plots
plots <- create_specific_plots(results)
print(plots$combined_plot)

# If you want to see individual plots:
# print(plots$niche_plot)
# print(plots$ri_biomass_plot)
# print(plots$species_biomass_plot)
# print(plots$community_biomass_plot)

```
## Formula
For species i at time t:
Biomass(i,t) = Biomass(i,t-1) + New_Biomass(i,t)

Where New_Biomass(i,t) = actual_N_uptake + actual_P_uptake

And the actual resource uptake is calculated as:
actual_N_uptake = k(i,t) * xn(i) * 5 * limitation_factor
actual_P_uptake = k(i,t) * yp(i) * 5 * limitation_factor

Where:

k(i,t) = k_max(i) * sigmoid_growth_factor
k_max(i) = sqrt(xn(i)^2 + yp(i)^2) * ni(i)
ni(i) depends on m and ri:
When m = 0: ni = 1
When m = 1: ni = ri
When m = -1: ni = 1/ri
Variables meaning:

i: species index
t: time point
xn: nitrogen uptake coefficient
yp: phosphorus uptake coefficient
k_max: maximum resource acquisition rate
ni: niche interaction effect
ri: sum of niche overlap + 1
m: interaction type (-1: competition, 0: neutral, 1: facilitation)
limitation_factor: resource availability constraint

## While m = 0 or 1 or -1
```{r}
# Function to run multiple simulations and collect results
run_multiple_simulations <- function(r, t, m, n_simulations = 20) {
  # Store results
  biomass_curves <- array(0, dim = c(t, n_simulations, 1))
  final_biomass <- numeric(n_simulations)
  
  for(sim in 1:n_simulations) {
    results <- resource_model(r, t, m)
    biomass_curves[,sim,1] <- results$total_biomass
    final_biomass[sim] <- results$total_biomass[t]
  }
  
  return(list(
    biomass_curves = biomass_curves,
    final_biomass = final_biomass
  ))
}

# Parameters
t <- 2000
richness_levels <- c(1, 4, 8, 12)
m_values <- c(-1, 0, 1)
n_sims <- 20

# Store all results
all_results <- list()
for(m in m_values) {
  all_results[[paste0("m", m)]] <- list()
  for(r in richness_levels) {
    all_results[[paste0("m", m)]][[paste0("r", r)]] <- run_multiple_simulations(r, t, m, n_sims)
  }
}

# Create visualization
create_scenario_plots <- function(results, m_value, scenario_name) {
  # Prepare data for average biomass curves
  curves_data <- data.frame()
  final_biomass_data <- data.frame()
  
  for(r in richness_levels) {
    # Average biomass curves
    avg_curve <- apply(results[[paste0("m", m_value)]][[paste0("r", r)]]$biomass_curves[,,1], 1, mean)
    curves_data <- rbind(curves_data, 
                        data.frame(
                          Time = 1:t,
                          Biomass = avg_curve,
                          Richness = factor(r)
                        ))
    
    # Final biomass for boxplot
    final_biomass_data <- rbind(final_biomass_data,
                               data.frame(
                                 Richness = factor(r),
                                 FinalBiomass = results[[paste0("m", m_value)]][[paste0("r", r)]]$final_biomass
                               ))
  }
  
  # Create average biomass curves plot
  p1 <- ggplot(curves_data, aes(x = Time, y = Biomass, color = Richness)) +
    geom_line(size = 1) +
    labs(title = paste("Average Biomass Curves -", scenario_name),
         x = "Time",
         y = "Total Biomass") +
    theme_minimal() +
    scale_color_brewer(palette = "Set1")
  
  # Create boxplot
  p2 <- ggplot(final_biomass_data, aes(x = Richness, y = FinalBiomass, fill = Richness)) +
    geom_boxplot() +
    labs(title = paste("Final Biomass Distribution -", scenario_name),
         x = "Species Richness",
         y = "Final Biomass") +
    theme_minimal() +
    scale_fill_brewer(palette = "Set1")
  
  return(list(curves_plot = p1, boxplot = p2))
}

# Generate plots for each scenario
competition_plots <- create_scenario_plots(all_results, -1, "Competition (m=-1)")
neutral_plots <- create_scenario_plots(all_results, 0, "Neutral (m=0)")
facilitation_plots <- create_scenario_plots(all_results, 1, "Facilitation (m=1)")

# Combine all plots
combined_plots <- (competition_plots$curves_plot + competition_plots$boxplot) /
                 (neutral_plots$curves_plot + neutral_plots$boxplot) /
                 (facilitation_plots$curves_plot + facilitation_plots$boxplot) +
  plot_layout(heights = c(1, 1, 1)) +
  plot_annotation(title = "Community Dynamics Across Different Scenarios",
                 theme = theme_minimal())

# Display the combined plot
print(combined_plots)

# Save the plot if needed
# ggsave("community_dynamics.png", combined_plots, width = 12, height = 15)


```

## np = n/p
```{r}
# Modified resource model with linked niche parameters
resource_model_linked <- function(r, t, m=1) {
  # Initialize resource pools
  initial_N <- 100
  initial_P <- 100
  N_pool <- initial_N
  P_pool <- initial_P
  tr <- 12
  community_richness <- tr/r
  
  # Create species characteristics
  species_traits <- data.frame(
    species = 1:r,
    xn = runif(r, min=0.1, max=1),
    yp = runif(r, min=0.1, max=1),
    growth_time = runif(r, min=10, max=20)
  )
  
  # Calculate base kmax
  species_traits$k_max <- sqrt(species_traits$xn^2 + species_traits$yp^2)
  
  # Link niche parameters to resource utilization
  species_traits$np <- species_traits$xn / species_traits$yp * 50  # Scale to reasonable range
  species_traits$nb <- species_traits$k_max * 15  # Scale to reasonable range
  
  # Calculate niche overlap and modify kmax
  niche_results <- calculate_niche_interactions(species_traits)
  species_traits$ri <- niche_results$ri_values
  species_traits$ni <- sapply(species_traits$ri, calculate_ni, m=m)
  species_traits$k_max <- species_traits$k_max * species_traits$ni
  
  # Initialize tracking matrices
  biomass_time <- matrix(0, nrow=t, ncol=r)
  total_biomass <- numeric(t)
  N_available <- numeric(t)
  P_available <- numeric(t)
  
  # Simulation loop
  for(time in 1:t) {
    N_available[time] <- N_pool
    P_available[time] <- P_pool
    
    if(N_pool > 0 && P_pool > 0) {
      # Calculate current acquisition rates
      current_k <- sapply(1:r, function(i) {
        if(time < species_traits$growth_time[i]) {
          species_traits$k_max[i] * (1 / (1 + exp(-5 * (time/species_traits$growth_time[i] - 0.5))))
        } else {
          species_traits$k_max[i]
        }
      })
      
      for(i in 1:r) {
        # Calculate potential uptake
        N_uptake <- current_k[i] * species_traits$xn[i] * community_richness
        P_uptake <- current_k[i] * species_traits$yp[i] * community_richness
        
        # Resource limitation
        N_limitation <- N_uptake/N_pool
        P_limitation <- P_uptake/P_pool
        limitation_factor <- min(1, 1/exp(max(N_limitation, P_limitation)))
        
        # Actual resource uptake
        actual_N_uptake <- N_uptake * limitation_factor
        actual_P_uptake <- P_uptake * limitation_factor
        
        # Update resource pools
        N_pool <- max(0, N_pool - actual_N_uptake)
        P_pool <- max(0, P_pool - actual_P_uptake)
        
        # Calculate new biomass increment
        new_biomass <- actual_N_uptake + actual_P_uptake
        
        # Accumulate biomass
        if(time == 1) {
          biomass_time[time, i] <- new_biomass
        } else {
          biomass_time[time, i] <- biomass_time[time-1, i] + new_biomass
        }
      }
    } else {
      if(time > 1) {
        biomass_time[time,] <- biomass_time[time-1,]
      }
    }
    
    total_biomass[time] <- sum(biomass_time[time,])
  }
  
  return(list(
    species_traits = species_traits,
    biomass_time = biomass_time,
    total_biomass = total_biomass,
    N_available = N_available,
    P_available = P_available
  ))
}

# Run simulations and create visualizations
create_comparison_plots <- function(n_sims=20, r=12, t=200) {
  # Store results
  linked_results <- list()
  independent_results <- list()
  
  for(i in 1:n_sims) {
    # set.seed(i)  # For reproducibility
    
    # Run both models
    linked <- resource_model_linked(r, t)
    independent <- resource_model(r, t, m=1)
    
    # Store results
    linked_results[[i]] <- linked
    independent_results[[i]] <- independent
  }
  
  # Prepare data for plotting
  # Figure 1: xn/yp vs final biomass
  xn_yp_data <- data.frame()
  for(i in 1:n_sims) {
    final_biomass <- linked_results[[i]]$biomass_time[t,]
    xn_yp_ratio <- linked_results[[i]]$species_traits$xn / 
                   linked_results[[i]]$species_traits$yp
    
    xn_yp_data <- rbind(xn_yp_data,
                        data.frame(
                          ratio = xn_yp_ratio,
                          biomass = final_biomass,
                          simulation = i
                        ))
  }
  
  fig1 <- ggplot(xn_yp_data, aes(x = ratio, y = biomass)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = TRUE) +
    labs(title = "Resource Utilization Ratio vs Final Biomass",
         x = "Resource Utilization Ratio (xn/yp)",
         y = "Final Biomass") +
    theme_minimal()
  
  # Figure 2: k vs final biomass
  k_data <- data.frame()
  for(i in 1:n_sims) {
    final_biomass <- linked_results[[i]]$biomass_time[t,]
    k_values <- linked_results[[i]]$species_traits$k_max
    
    k_data <- rbind(k_data,
                    data.frame(
                      k = k_values,
                      biomass = final_biomass,
                      simulation = i
                    ))
  }
  
  fig2 <- ggplot(k_data, aes(x = k, y = biomass)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = TRUE) +
    labs(title = "Resource Acquisition Rate vs Final Biomass",
         x = "Resource Acquisition Rate (k)",
         y = "Final Biomass") +
    theme_minimal()
  
  # Figure 3: Biomass curves for linked model
  linked_avg_community <- matrix(0, nrow=t, ncol=1)
  linked_avg_species <- matrix(0, nrow=t, ncol=r)
  
  for(i in 1:n_sims) {
    linked_avg_community[,1] <- linked_avg_community[,1] + 
                               linked_results[[i]]$total_biomass
    linked_avg_species <- linked_avg_species + linked_results[[i]]$biomass_time
  }
  
  linked_avg_community <- linked_avg_community / n_sims
  linked_avg_species <- linked_avg_species / n_sims
  
  linked_data <- data.frame(
    Time = rep(1:t, r+1),
    Biomass = c(as.vector(linked_avg_species), linked_avg_community),
    Type = factor(rep(c(paste("Species", 1:r), "Community"), each=t))
  )
  
  fig3 <- ggplot(linked_data, aes(x = Time, y = Biomass, color = Type)) +
    geom_line(size = 1) +
    labs(title = "Biomass Dynamics (Linked Niche Parameters)",
         x = "Time",
         y = "Biomass") +
    theme_minimal()
  
  # Figure 4: Biomass curves for independent model
  indep_avg_community <- matrix(0, nrow=t, ncol=1)
  indep_avg_species <- matrix(0, nrow=t, ncol=r)
  
  for(i in 1:n_sims) {
    indep_avg_community[,1] <- indep_avg_community[,1] + 
                              independent_results[[i]]$total_biomass
    indep_avg_species <- indep_avg_species + independent_results[[i]]$biomass_time
  }
  
  indep_avg_community <- indep_avg_community / n_sims
  indep_avg_species <- indep_avg_species / n_sims
  
  indep_data <- data.frame(
    Time = rep(1:t, r+1),
    Biomass = c(as.vector(indep_avg_species), indep_avg_community),
    Type = factor(rep(c(paste("Species", 1:r), "Community"), each=t))
  )
  
  fig4 <- ggplot(indep_data, aes(x = Time, y = Biomass, color = Type)) +
    geom_line(size = 1) +
    labs(title = "Biomass Dynamics (Independent Niche Parameters)",
         x = "Time",
         y = "Biomass") +
    theme_minimal()
  
  # Combine all plots
  combined_plot <- (fig1 + fig2) / (fig3 + fig4) +
    plot_annotation(title = "Community Dynamics Comparison",
                   theme = theme_minimal())
  
  return(combined_plot)
}

# Run the analysis and display results
plots <- create_comparison_plots()
print(plots)
```